---
title: "Try to Modelling"
author: "Giulia Grossi"
date: "26 settembre 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Packages upload
```{r}
library(readr)
library(readxl)
library(dplyr)
library(ISLR)
library(tidyr)
```

# 1. Merging ABS dataset (for BEC)
Informatio for Bec on merging datasets of ABS _demo and socio 
```{r}
##### merging abs_datasets #####
setwd("~/UTS/STAT. THINKING/Groupwork_vaccination/abs datasets")
abs_pop=read.csv("./ABS_REGIONAL_pop.csv", header=TRUE)
abs_famcom=read.csv("./ABS_REGIONAL_family&community.csv", header=TRUE)
abs_edu=read.csv("./ABS_REGIONAL_educ&employment.csv", header=TRUE)
abs_eco=read.csv("./ABS_REGIONAL_economy&income.csv", header=TRUE)



datasheet_report <- read_excel("~/UTS/STAT. THINKING/Groupwork_vaccination/abs datasets/datasheet-report-hc42_Loki.xlsx", 
                               sheet = "TAB 3", col_types = c("text","text", "text", "text", "blank",
                                                              "numeric", "numeric", "numeric", 
                                                                "numeric", "numeric"))

glimpse(datasheet_report)

abs_data <- rbind(abs_eco,abs_edu,abs_famcom,abs_pop)
#delete insignificant columns:
abs_data <- abs_data %>% 
  select(-c("Geography.Level","FREQUENCY","TIME","Flag.Codes","Flags")) 

#rename columns for merging
names(abs_data)[4] <- "SA3code"
names(datasheet_report)[2] <- "SA3code"
names(datasheet_report)[4] <- "Time"

#merge by SA3 and Year
DF = merge(abs_data,datasheet_report, by= c("SA3code", "Time"))
View(DF)
unique(DF$Time)


write.csv(DF, "All abs by year.csv")
```


# 2. Work on ABS dataset to discover some models

Dataset re-merde with correction on "year" anomalies (before there were data referred to different years in the same row). 
Datasetname "All abs by year", 
```{r}
All_abs_by_year <- read_csv("~/UTS/STAT. THINKING/Groupwork_vaccination/2B_modelling/All abs by year.csv")
unique(All_abs_by_year$Data.item) 
```
There are 98 different feature under "Data.item" --> extractthe most interesting ones to create new indipendent variables.

Aim: extract socio-economic indicators to find possible correlations
```{r}
CLEAR_ABS <- spread(All_abs_by_year, Data.item, Value)

CLEAR_ABS <- CLEAR_ABS %>% 
  select(-c(1,20,21,42,57,86,88,91,18,19,34,38,39,28))
str(CLEAR_ABS)
```

### plot to fin possible interesting correlations
```{r}
attach(CLEAR_ABS)
plot(`Percent fully immunised (%)` ~ `Average number of all children in a family`)
plot(`Percent fully immunised (%)` ~ `Australian citizen (%)`)
plot(`Percent fully immunised (%)` ~ `Not an Australian citizen (%)`)

plot(`Percent fully immunised (%)` ~ `Did not go to school (%)`)
plot(`Percent fully immunised (%)`~ `Completed Year 10 or equivalent (%)`)
plot(`Percent fully immunised (%)`~ `Internet accessed from dwelling (%)`)
plot(`Percent fully immunised (%)`~ `Males - Total (no.)`)
plot(`Percent fully immunised (%)`~ `Managers (%)`)
plot(`Percent fully immunised (%)`~ `Married (%)`)  ##NB low values
plot(`Percent fully immunised (%)`~ `Mean employee income ($)`)  #NB low values!
plot(`Percent fully immunised (%)`~ `Median Age - Females (years)`)  #NB low values!
plot(`Percent fully immunised (%)`~ `Median Age - Males (years)`)  
plot(`Percent fully immunised (%)`~ `Persons earning $500-$999 per week (%)`)  
plot(`Percent fully immunised (%)`~ `Persons earning $1-$499 per week (%)`)  
plot(`Percent fully immunised (%)`~ `Persons earning $3000 or more per week (%)`)  
plot(`Percent fully immunised (%)`~ `Total income (excl. Government pensions and allowances) ($)`) #NB low values
plot(`Percent fully immunised (%)`~ `Total persons employed (no.)`) #NB low 
plot(`Percent fully immunised (%)`~ `Unemployment rate (%)`) #NB low 

detach(CLEAR_ABS)
```

###Prepare data: create a binary %immunisation as response vabiable
```{r}
CLEAR_ABS$Immu01 <- ifelse(CLEAR_ABS$`Percent fully immunised (%)` >= 95.0, 1,0)
summary
CLEAR_ABS$Immu01 <- factor(CLEAR_ABS$Immu01)
table(CLEAR_ABS$Immu01)  #unbalanced datasets --> 11.5% in target

#make "Time" a factor: 
CLEAR_ABS$Time <- factor(CLEAR_ABS$Time, levels= c(2011:2016))
str(CLEAR_ABS$Time)

```

Analyse boxplots using %Immu01 as output variable.
I decided to focus on these soco-ec.indicators:
```{r}
boxplot_immu <- function(CLEAR_ABS,variable, label) {
  CLEAR_ABS %>% ggplot(aes(x = Immu01, y=variable, fill=Immu01)) + geom_boxplot() + labs(y=label)
}

boxplot_immu(CLEAR_ABS,CLEAR_ABS$`Unemployment rate (%)`, "Unemployment rate")
boxplot_immu(CLEAR_ABS,CLEAR_ABS$`Unemployment rate (%)`, "Unemployment rate")
boxplot_immu(CLEAR_ABS,CLEAR_ABS$`Total persons employed (no.)`, "Total Employee")
boxplot_immu(CLEAR_ABS,CLEAR_ABS$`Total income (excl. Government pensions and allowances) ($)`, "Total income")
boxplot_immu(CLEAR_ABS,CLEAR_ABS$`Median Age - Females (years)`, "female median age")
boxplot_immu(CLEAR_ABS,CLEAR_ABS$`Married (%)`, "marriage %")

```
Result: nothing particularly interesting, a lot of NA to manage and clean.

### Logistic model
Not significant
```{r}
summary(glm(Immu01 ~ `Unemployment rate (%)`+
              `Age group` , data = CLEAR_ABS, family="binomial"))
```





# Investigate prediction on tax, phn, postcode, seifa score
Work on dataset "Immunization with everything taxation update"
```{r}
ethan <- read_csv("C:/giulia/Vaccination/cleaned_data/immunization_with_everything_taxation_update.csv")
View(ethan)
```

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
