---
title: "Merge Taxation with Immunisation datasets"
output: html_document
author: Ethan- Quoc Do
---


```{r setup, include=FALSE}
#Load all neccessary library for this merge
library(dplyr)
library(readr)
library(tidyverse)
library(readxl)
options(scipen=999)

```


## Import The immunisation dataset 

This dataset is already included percent fo immunisation, PHN, SEIFA, electro.

```{r}
immu_clean_raw<-read.csv("immunization_with_everything.csv")
summary(immu_clean_raw) #provide both summary and str to check
str(immu_clean_raw) 
#quick note,there are alot of NA in SEIFA c& Electro data. But it will not affect our merge
```

## Import Taxation

#### 2016
In the purpose of this merge, we only need to use year and postcode as two `keys` to merge them together. Taxation datasets are provided years by years. This will be an advantage for merging because it will take lesser resoures to combine. Furthermore, postcode is changing over the course of year. Thus, It better to combine them seperately years by years.

```{r}
######Import Taxation in 2016#####
tax_2016<-read_xlsx("2016.xlsx", sheet="Individuals Table 6B")
#Select only Postcode, Taxable Income and Number of individuals
str(tax_2016)
```

<br>

Column are terribily naming after importing in R. Let's make it easier to us by changing the name
```{r}
#rename them first for easy to merge
tax_2016<-tax_2016%>%
  rename("state"= "State/ Territory1")%>%
  rename("num_individuals"= "Number of individuals no.")%>%
rename("total_tax"= "Taxable income or loss $")%>%
rename("postcode"="Postcode")
```

```{r}
#remove any " , " in these two columns and convert to number
tax_2016$total_tax<-as.numeric(tax_2016$total_tax)
tax_2016$total_tax<-gsub(",","",tax_2016$total_tax)
tax_2016$num_individuals<-as.numeric(tax_2016$num_individuals)
tax_2016$total_tax<-as.numeric(tax_2016$total_tax)
```
<br>
We only want the number of individuals and total tax per post

```{r}
#filter out uneccessary column. 
#Select only Postcode, Taxable Income and Number of individuals
tax_2016<-tax_2016%>%
  select(state,postcode,num_individuals,total_tax)
```

```{r}
# create new column: mean of tax
tax_2016_final<-tax_2016%>%
  group_by(postcode)%>%
mutate(mean_tax= ceiling(total_tax/num_individuals))%>%
  mutate(year="2016")
```

`year` and `postcode` both integer. Let's cover them
```{r}
#convert integert to have the same column character
tax_2016_final$year<-as.integer(tax_2016_final$year)
tax_2016_final$postcode<-as.integer(tax_2016_final$postcode)
```
```{r}
#This introduce some NAs, which include non postcode one, let's remove them
any(is.na(tax_2016_final$postcode))
sum(is.na(tax_2016_final$postcode))
#Remove the record that doesn't define postcode but in name such as NSW and other.
tax_2015_final<-na.omit(tax_2016_final)
```
```{r}
# Create Immu data for only 2016
immu_2016<-immu_clean_raw%>%
     filter(year==2016)

#Join data with immuni_clean
combine.2016<-left_join(immu_2016,tax_2016_final,by=c("year","postcode"))
```

We finish with 2016. The later part will be the same for 2015, 2013, 2011, except 2014 and 2012. Because these datasets do not have total tax but slipping into "non taxable" and "taxable", which we need to calculate the sum before merge.


#### Taxataion from 2011 to 2015
```{r}

#####Import Taxation in 2015######
tax_2015<-read_xlsx("2015.xlsx", sheet="Individuals Table 6B")
names(tax_2015)
summary(tax_2015$postcode)
#rename them first for easy to merge
tax_2015<-tax_2015%>%
  rename("state"= "State/ Territory1")%>%
  rename("num_individuals"= "Number of individuals\r\nno.")%>%
  rename("total_tax"= "Taxable income or loss3\r\n$")%>%
  rename("postcode"="Postcode")

#remove any " , " in these two columns and convert to number
tax_2015$total_tax<-as.numeric(tax_2015$total_tax)
tax_2015$total_tax<-gsub(",","",tax_2015$total_tax)
tax_2015$num_individuals<-as.numeric(tax_2015$num_individuals)
tax_2015$total_tax<-as.numeric(tax_2015$total_tax)



#filter out uneccessary column
tax_2015<-tax_2015%>%
  select(state,postcode,num_individuals,total_tax)

# create new column for mean values and year
tax_2015_final<-tax_2015%>%
  group_by(postcode)%>%
  mutate(mean_tax= ceiling(total_tax/num_individuals))%>%
  mutate(year="2015")

#convert integert to have the same column character
tax_2015_final$year<-as.integer(tax_2015_final$year)
tax_2015_final$postcode<-as.integer(tax_2015_final$postcode)
#This introduce some NAs, which include non postcode one, let's remove them
any(is.na(tax_2015_final$postcode))
sum(is.na(tax_2015_final$postcode))
tax_2015_final<-na.omit(tax_2015_final)

#create immuni data for 2015 only
immu_2015<-immu_clean_raw%>%
  filter(year==2015)

#Join data immuni_2015 with tax_2015_final
combine.2015<-left_join(immu_2015,tax_2015_final,by=c("year","postcode"))
#####END 2015 ####


#### Import taxtation 2014 #####
tax_2014<-read_xlsx("2014.xlsx", sheet="Table 6")
names(tax_2014)
str(tax_2014$postcode)
summary(tax_2014$postcode)
glimpse(tax_2014$`Taxable Status`)
#rename them first for easy to merge
tax_2014<-tax_2014%>%
  rename("state"= "State/Territory1")%>%
  rename("num_individuals"= "Number of individuals")%>%
  rename("total_tax"= "Taxable income or loss3\r\n$")%>%
  rename("postcode"="Postcode")%>%
rename("taxable_stat"="Taxable Status")
#remove any " , " in these two columns and convert to number
tax_2014$total_tax<-as.numeric(tax_2014$total_tax)
tax_2014$total_tax<-gsub(",","",tax_2014$total_tax)
tax_2014$num_individuals<-as.numeric(tax_2014$num_individuals)
tax_2014$total_tax<-as.numeric(tax_2014$total_tax)



#filter out uneccessary column
tax_2014<-tax_2014%>%
  select(state,postcode,num_individuals,total_tax)
#extra step because 2014 doesn't combine the data Taxable and Non Tax
tax_2014_test<-tax_2014 %>% 
  group_by(postcode,state) %>% 
  summarise(total_tax = sum(total_tax),num_individuals = sum(num_individuals))

# create new column for mean values and year
tax_2014_final<-tax_2014_test%>%
  group_by(postcode)%>%
  mutate(mean_tax= ceiling(total_tax/num_individuals))%>%
  mutate(year="2014")




#convert integert to have the same column character
tax_2014_final$year<-as.integer(tax_2014_final$year)
tax_2014_final$postcode<-as.integer(tax_2014_final$postcode)
#This introduce some NAs, which include non postcode one, let's remove them
any(is.na(tax_2014_final$postcode))
sum(is.na(tax_2014_final$postcode))
tax_2014_final<-na.omit(tax_2014_final)

#create immuni data for 2014 only
immu_2014<-immu_clean_raw%>%
  filter(year==2014)

#Join data immuni_2014 with tax_2014_final
combine.2014<-left_join(immu_2014,tax_2014_final,by=c("year","postcode"))
#####END 2014 ####
#### Import taxtation 2013 #####
tax_2013<-read_xlsx("2013.xlsx", sheet="Postcode only")
names(tax_2013)
str(tax_2013$postcode)
summary(tax_2013$postcode)
glimpse(tax_2013$`Taxable Status`)
#rename them first for easy to merge
tax_2013<-tax_2013%>%
  rename("state"= "State/Territory1")%>%
  rename("num_individuals"= "Number of individuals\r\nno.")%>%
  rename("total_tax"= "Taxable income or loss\r\n$")%>%
  rename("postcode"="Postcode")

#remove any " , " in these two columns and convert to number
tax_2013$total_tax<-as.numeric(tax_2013$total_tax)
tax_2013$total_tax<-gsub(",","",tax_2013$total_tax)
tax_2013$num_individuals<-as.numeric(tax_2013$num_individuals)
tax_2013$total_tax<-as.numeric(tax_2013$total_tax)



#filter out uneccessary column
tax_2013<-tax_2013%>%
  select(state,postcode,num_individuals,total_tax)

# create new column for mean values and year
tax_2013_final<-tax_2013%>%
  group_by(postcode)%>%
  mutate(mean_tax= ceiling(total_tax/num_individuals))%>%
  mutate(year="2013")

#convert integert to have the same column character
tax_2013_final$year<-as.integer(tax_2013_final$year)
tax_2013_final$postcode<-as.integer(tax_2013_final$postcode)
#This introduce some NAs, which include non postcode one, let's remove them
any(is.na(tax_2013_final$postcode))
sum(is.na(tax_2013_final$postcode))
tax_2013_final<-na.omit(tax_2013_final)

#create immuni data for 2013 only
immu_2013<-immu_clean_raw%>%
  filter(year==2013)

#Join data immuni_2013 with tax_2013_final
combine.2013<-left_join(immu_2013,tax_2013_final,by=c("year","postcode"))
#####END 2013 ####
#### Import taxtation 2012 #####
tax_2012<-read_xlsx("2012.xlsx", sheet="Individuals Tax Table 6")
names(tax_2012)
str(tax_2012$postcode)
summary(tax_2012$postcode)
glimpse(tax_2012$`Taxable Status`)
#rename them first for easy to merge
tax_2012<-tax_2012%>%
  rename("state"= "State/Territory1")%>%
  rename("num_individuals"= "Number of individuals\r\nno.")%>%
  rename("total_tax"= "Taxable income or loss3\r\n$")%>%
  rename("postcode"="Postcode")

#remove any " , " in these two columns and convert to number
tax_2012$total_tax<-as.numeric(tax_2012$total_tax)
tax_2012$total_tax<-gsub(",","",tax_2012$total_tax)
tax_2012$num_individuals<-as.numeric(tax_2012$num_individuals)
tax_2012$total_tax<-as.numeric(tax_2012$total_tax)



#filter out uneccessary column
tax_2012<-tax_2012%>%
  select(state,postcode,num_individuals,total_tax)


#extra step because 2014 doesn't combine the data
tax_2012_test<-tax_2012 %>% 
  group_by(postcode,state) %>% 
  summarise(total_tax = sum(total_tax),num_individuals = sum(num_individuals))

# create new column for mean values and year
tax_2012_final<-tax_2012_test%>%
  group_by(postcode)%>%
  mutate(mean_tax= ceiling(total_tax/num_individuals))%>%
  mutate(year="2012")

#convert integert to have the same column character
tax_2012_final$year<-as.integer(tax_2012_final$year)
tax_2012_final$postcode<-as.integer(tax_2012_final$postcode)
#This introduce some NAs, which include non postcode one, let's remove them
any(is.na(tax_2012_final$postcode))
sum(is.na(tax_2012_final$postcode))
tax_2012_final<-na.omit(tax_2012_final)

#create immuni data for 2012 only
immu_2012<-immu_clean_raw%>%
  filter(year==2012)

#Join data immuni_2012 with tax_2012_final
combine.2012<-left_join(immu_2012,tax_2012_final,by=c("year","postcode"))

#####END 2012 ####

#### Import taxtation 2011 #####
tax_2011<-read_xlsx("2011.XLSX", sheet="Individuals Tax Table 6a")
names(tax_2011)
str(tax_2011$postcode)
summary(tax_2011$postcode)
glimpse(tax_2011$`Taxable Status`)
#rename them first for easy to merge
tax_2011<-tax_2011%>%
  rename("state"= "State/Territory")%>%
  rename("num_individuals"= "Number of individuals")%>%
  rename("total_tax"= "Taxable income or loss")%>%
  rename("postcode"="Postcode")

#remove any " , " in these two columns and convert to number
tax_2011$total_tax<-as.numeric(tax_2011$total_tax)
tax_2011$total_tax<-gsub(",","",tax_2011$total_tax)
tax_2011$num_individuals<-as.numeric(tax_2011$num_individuals)
tax_2011$total_tax<-as.numeric(tax_2011$total_tax)



#filter out uneccessary column
tax_2011<-tax_2011%>%
  select(state,postcode,num_individuals,total_tax)

# create new column for mean values and year
tax_2011_final<-tax_2011%>%
  group_by(postcode)%>%
  mutate(mean_tax= ceiling(total_tax/num_individuals))%>%
  mutate(year="2011")

#convert integert to have the same column character
tax_2011_final$year<-as.integer(tax_2011_final$year)
tax_2011_final$postcode<-as.integer(tax_2011_final$postcode)
#This introduce some NAs, which include non postcode one, let's remove them
any(is.na(tax_2011_final$postcode))
sum(is.na(tax_2011_final$postcode))
tax_2011_final<-na.omit(tax_2011_final)

#create immuni data for 2011 only
immu_2011<-immu_clean_raw%>%
  filter(year==2011)

#Join data immuni_2011 with tax_2011_final
combine.2011<-left_join(immu_2011,tax_2011_final,by=c("year","postcode"))
#####END 2011 ####
```


## Merge all the year:

```{r}
##let's combine them
test1 <- rbind(combine.2016,combine.2015)
test2 <- rbind(test1,combine.2014)
test3 <- rbind(test2,combine.2013)
test4 <-rbind(test3,combine.2012)
test5 <- rbind(test4,combine.2011)
```

After this, writing it out and share to the group

```{r}
write.csv(test5,"immunization_with_everything_taxation_update.csv")
```


Thanks for reading. Peace!!
