---
title: "EDA_Vaccinations"
author: "Alex Brooks"
date: "19 September 2018"
output: html_document
---
### I would like to explore the bottom postcodes for low immun % coverage and the top postcodes for immun % coverage

This task was extremely hard, as I could only identify bottom 20 postcodes and top 20 postcodes, but due to duplicates I could then only explore the bottom class of immunisation % coverage (<70% and >95%) and the top class.  


```{r}
#load libraries
library(readr)
library(ggplot2)
library(dplyr)
library(Amelia)
library(reshape2)
```

## Immport and clean data

I used the merge file containing, PHN, postcode, immun%coverage, SEIFA data and electoral data. (IE, there were no demographics or taxable income data in this file)
```{r}
#import newly merged dataset
immunisations_merged <- read.csv("../cleaned_data/immunization_with_everything.csv")
str(immunisations_merged)
```

```{r}
#check to see how it really looks - it's a mix of long and wide data
summary(immunisations_merged)
```

1. Reduced columns names
```{r}
# Clean data using function 

# Parameters
# * imm_merged - the source data frame we want to clean

# Returns - a data frame

data <- immunisations_merged %>%
    select('state', 'postcode', 'year', 'age',  'pc_immun', 'caution', 'pc_immun_class','PHN_code', 'PHN_number','Usual.resident.population','Index.type','Score', 'Electoral.division', 'PartyNm2016', 'PartyNm2010', 'PartyNm2013')

data$age <- as.factor(data$age)

str(data)

```



```{r}
#Manipulate the data to get distinct PC Immun groups

pc_immun_groups <- data %>%
    group_by(pc_immun, pc_immun_class) %>%
    summarize(count = n())

pc_immun_groups

```

```{r}
#Manipulate the data to get distinct age groups

age_groups <- data %>%
    group_by(age) %>%
    summarize(count = n())

age_groups

```
##Top % coverage areas versus bottom % coverage areas 
2. I want to see the bottom postcodes for immunisation coverage and the top postcodes for immunisation coverage


```{r}
##creating bottom % immunisation coverage in 2016 to group, summarise and arrange
low_pc_immun <- data %>%
  filter(pc_immun_class < 3, pc_immun_class != 0, Index.type == "Index of Relative Socio-economic Advantage and Disadvantage", year == 2016) %>%
  group_by(postcode, PHN_code, state, pc_immun, pc_immun_class, Usual.resident.population) %>%
  summarize(count = n()) %>%
  arrange(pc_immun_class, desc(Usual.resident.population))

low_pc_immun
View(low_pc_immun)
```


```{r}
bottom_20_postocdes <- head(low_pc_immun, 20)

bottom_20_postocdes
#2016 results
```

```{r}
#order and rank the top 20 in 2016
pc_immun_ordered <- data %>%
  group_by(postcode, PHN_code, state, pc_immun, pc_immun_class, Usual.resident.population) %>%
  filter(pc_immun_class != 0, Index.type == "Index of Relative Socio-economic Advantage and Disadvantage", year == 2016) %>%
  summarize(count = n()) %>%
  arrange(desc(pc_immun_class), desc(Usual.resident.population))


pc_immun_ordered
View(pc_immun_ordered)
```

```{r}
top_20_postocdes <- head(pc_immun_ordered, 20)

top_20_postocdes
```

```{r}
bottom_20_postocdes <- tail(pc_immun_ordered, 20)

arrange(bottom_20_postocdes, pc_immun_class, desc(Usual.resident.population))
```



3. I want to know the mean immunisation % coverage rate across Australia in 2011 - but I also want to know what the mean immunisation% coverage rate is for each state.
```{r}
#filter and explore the 2011 data
mean_immun_rate_aust_2011 <- data %>%
    filter(year == 2011, pc_immun_class != 0)

mean(mean_immun_rate_aust_2011$pc_immun_class)

```

```{r}
#filter, group and summarise the 2011 data by state
mean_immun_rate_state_2011 <- data %>%
    filter(year == 2011) %>%
    group_by(state) %>%
    summarize(
        mean = ifelse(sum(pc_immun_class[pc_immun_class != 0]) > 0, mean(pc_immun_class[pc_immun_class != 0]), 0),
        rounded_mean = round(as.numeric(mean)),
        pc_immun = pc_immun_groups$pc_immun[pc_immun_groups$pc_immun_class == rounded_mean],
        total_rows = n(),
        num_of_non_NP = sum(pc_immun_class != 0),
        num_of_NP = sum(pc_immun_class == 0),
        proportion_of_NP = (num_of_NP/n()))

mean_immun_rate_state_2011

```


4. I want to know the mean immunisation % coverage rate across Australia in 2016 - and I want to know what the mean is for each state (even those weird doubly states we have)

```{r}
mean_immun_rate_aust_2016 <- data %>%
    filter(year == 2016, pc_immun_class != 0)

mean(mean_immun_rate_aust_2016$pc_immun_class)

```

```{r}
#filter, group and summarise the 2016 data by state
mean_immun_rate_state_2016 <- data %>%
    filter(year == 2016) %>%
    group_by(state) %>%
    summarize(
        mean = ifelse(sum(pc_immun_class[pc_immun_class != 0]) > 0, mean(pc_immun_class[pc_immun_class != 0]), 0),
        rounded_mean = round(as.numeric(mean)),
        pc_immun = pc_immun_groups$pc_immun[pc_immun_groups$pc_immun_class == rounded_mean],
        total_rows = n(),
        num_of_non_NP = sum(pc_immun_class != 0),
        num_of_NP = sum(pc_immun_class == 0),
        proportion_of_NP = (num_of_NP/n()))


mean_immun_rate_state_2016
View(mean_immun_rate_state_2016)
```

5. I want to see meaningful electorate and party in power for 2011 and 2016 - I don't know whether this would just be a smaller dataset or if we can somehow plot it???

```{r}

#Get info by electorate for 2016
electorates <- data %>%
    filter(Index.type == "Index of Relative Socio-economic Advantage and Disadvantage") %>%
    group_by(Electoral.division, PartyNm2016, year) %>%
    summarize(
        mean = ifelse(sum(pc_immun_class[pc_immun_class != 0]) > 0, mean(pc_immun_class[pc_immun_class != 0]), 0),
        rounded_mean = round(as.numeric(mean)),
        pc_immun = pc_immun_groups$pc_immun[pc_immun_groups$pc_immun_class == rounded_mean],
        total_rows = n(),
        num_of_non_NP = sum(pc_immun_class != 0),
        num_of_NP = sum(pc_immun_class == 0),
        proportion_of_NP = (num_of_NP/n())) %>%
    arrange(mean)

```

Get the electorates with the lowest 10 mean pc_immun_class scores for 2016
```{r}
lowest_pc_immun <- head(filter(electorates, year == 2016), 10)

lowest_pc_immun
```

Get those electorate names
```{r}
lowest_pc_immun_names <- select(lowest_pc_immun, Electoral.division, PartyNm2016)
lowest_pc_immun_names <- as.character(lowest_pc_immun$Electoral.division)

```

Let's look at postcodes in the electorates with lowest mean pc_immun_class scores

```{r}
#Get postciode level data for the lowest10 electorates in 2016
lowest_electorates_2016 <- data %>%
    filter(Electoral.division %in% lowest_pc_immun_names) %>%
    select(state, year, Index.type, Electoral.division, postcode, age, pc_immun, pc_immun_class, PHN_code, Score)
```

```{r fig.height=10, fig.width=12}
ggplot(filter(lowest_electorates_2016, Index.type == "Index of Relative Socio-economic Disadvantage"), aes(age)) +
  geom_point(aes(y = pc_immun, size=Score, color = age), na.rm = TRUE) +
  facet_wrap(~ year, ncol = 3, nrow = 5) +
  labs(title="Relative Socio-economic Disadvantage Scores", subtitle = "Bottom 10 electorates in 2016: a retrospective")

```

```{r fig.height=10, fig.width=12}
#OK. Let's look at postcodes in those electorates
#Get info by electorate
ggplot(filter(lowest_electorates_2016, Index.type == "Index of Relative Socio-economic Advantage and Disadvantage"), aes(age)) +
  geom_point(aes(y = pc_immun, size=Score, color = age), na.rm = TRUE) +
  facet_wrap(~ year, ncol = 3) +
  labs(title="Relative Socio-economic Advantage and Disadvantage Scores", subtitle = "Bottom 10 electorates in 2016: a retrospective")

```

```{r fig.height=10, fig.width=12}
#OK. Let's look at postcodes in those electorates
#Get info by electorate
ggplot(filter(lowest_electorates_2016, Index.type == "Index of Economic Resources"), aes(age)) +
  geom_point(aes(y = pc_immun, size=Score, color = age), na.rm = TRUE) +
  facet_wrap(~ year, ncol = 3) +
  labs(title="Economic Resources Scores", subtitle = "Bottom 10 electorates in 2016: a retrospective")

```

```{r fig.height=10, fig.width=12}
#OK. Let's look at postcodes in those electorates
#Get info by electorate
ggplot(filter(lowest_electorates_2016, Index.type == "Index of Education and Occupation"), aes(age)) +
  geom_point(aes(y = pc_immun, size=Score, color = age), na.rm = TRUE) +
  facet_wrap(~ year, ncol = 3) +
  labs(title="Education and Occupation Scores", subtitle = "Bottom 10 electorates in 2016: a retrospective")

```

```{r}
highest_pc_immun <- tail(filter(electorates, year == 2016), 10)

highest_pc_immun
```

```{r}
highest_pc_immun_names <- select(highest_pc_immun, Electoral.division, PartyNm2016)
highest_pc_immun_names <- as.character(highest_pc_immun$Electoral.division)
```

```{r}
#OK. Let's look at postcodes in those electorates
#Get info by electorate
#Get postcode level data for the highest 10 electorates in 2016
highest_electorates_2016 <- data %>%
    filter(Electoral.division %in% highest_pc_immun_names) %>%
    select(state, year, Electoral.division, Index.type, postcode, age, pc_immun, pc_immun_class, PHN_code, Score)

```


```{r fig.height=10, fig.width=12}
ggplot(filter(highest_electorates_2016, Index.type == "Index of Relative Socio-economic Disadvantage"), aes(age)) +
  geom_point(aes(y = pc_immun, size=Score, color = age), na.rm = TRUE) +
  facet_wrap(~ year, ncol = 3) +
  labs(title="Relative Socio-economic Disadvantage Score", subtitle = "Top 10 highest electorates in 2016: a retrospective")

```

```{r fig.height=10, fig.width=12}
#OK. Let's look at postcodes in those electorates
#Get info by electorate
ggplot(filter(highest_electorates_2016, Index.type == "Index of Relative Socio-economic Advantage and Disadvantage"), aes(age)) +
  geom_point(aes(y = pc_immun, size=Score, color = age), na.rm = TRUE) +
  facet_wrap(~ year, ncol = 3) +
  labs(title="Relative Socio-economic Advantage and Disadvantage Scores", subtitle = "Top 10 highest electorates in 2016: a retrospective")

```

```{r fig.height=10, fig.width=12}
#OK. Let's look at postcodes in those electorates
#Get info by electorate
ggplot(filter(highest_electorates_2016, Index.type == "Index of Economic Resources"), aes(age)) +
  geom_point(aes(y = pc_immun, size=Score, color = age), na.rm = TRUE) +
  facet_wrap(~ year, ncol = 3) +
  labs(title="Economic Resources Scores", subtitle = "Top 10 highest electorates in 2016: a retrospective")

```


```{r fig.height=10, fig.width=12}
ggplot(filter(highest_electorates_2016, Index.type == "Index of Education and Occupation"), aes(age)) +
  geom_point(aes(y = pc_immun, size = Score, color = age), na.rm = TRUE) +
  facet_wrap(~ year, ncol = 3) +
  labs(title="Education and Occupation Scores", subtitle = "Top 10 highest electorates in 2016: a retrospective")

```

```{r}
#Get info by 2016political party
party_2016 <- data %>%
    group_by(PartyNm2016) %>%
    summarize(
        mean = ifelse(sum(pc_immun_class[pc_immun_class != 0]) > 0, mean(pc_immun_class[pc_immun_class != 0]), 0),
        rounded_mean = round(as.numeric(mean)),
        pc_immun = pc_immun_groups$pc_immun[pc_immun_groups$pc_immun_class == rounded_mean],
        total_rows = n(),
        num_of_non_NP = sum(pc_immun_class != 0),
        num_of_NP = sum(pc_immun_class == 0),
        proportion_of_NP = (num_of_NP/n()))

party_2016
```
6. I want to somehow try to map or plot the SEIFA scores for the bottom 20 and top 20 postcodes for immunisation coverage - again, this might be too ambitious


## We know lots of postcodes are missing % immunisation coverage

##There seems to be some missing electoral data too

```{r}
#GETANT TO LOOK AT IT TO PRINT OUT FOR TODAY
#check for missing values

NP_per_state <- data %>%
  select(state, year, pc_immun_class) %>%
  filter(pc_immun_class == 0) %>%
  group_by(state, year) %>%
  summarize(count = n())

View(NP_per_state)
```

```{r}
#Plot NPs per state
ggplot(data = NP_per_state, mapping = aes(x = state,y = count, color = year))+
         geom_point() +
  labs(title="Not Published (NPs) counts by state per year", subtitle = "Note: 2011 had no NPs")
```

```{r}

missmap(data, main = "Missing values vs observed")
#read this to see if I should impute missing data https://www.linkedin.com/pulse/amelia-packager-missing-data-imputation-ramprakash-veluchamy/
```

```{r}
#try plotting postcode to immun_class over the years - not very visual
ggplot(data = data, mapping = aes(x = postcode,y = pc_immun_class, color = year))+
         geom_point()
```


```{r}
# Get % immunised counts per pc immun class for each year
counts_by_year_merged<-data %>%
  group_by(year, pc_immun_class) %>%
  summarize(count = n())

counts_by_year_merged$pc_immun = NA
View(counts_by_year_merged)
```

Now try plots


```{r}
#plot how % immunised has changed by category between 2011 to 2016 - I would like to not have the bottom counts so cluttered on this axis
ggplot(data = counts_by_year_merged) +
  geom_point(mapping = aes(x=count, y=year, color=pc_immun_class))+
  facet_wrap(~pc_immun, nrow = 2)

```


  